Staged Abstract Interpreters

### 2019-03-29 paper/40_absinterp.tex
evaluation

In this section, we simply let @R[T] = T@ (i.e., @NoRep@) to instantiate the
unstaged lattice type class. An simple instance of lattice we used is the power
set abstract domain (shown in Figure \ref{fig:powerset}), where the ordering is subset relation, the join is union
operation, and the meet is intersect operation.
Other lattices used in the rest of the paper, such as products and maps, can be
implemented similarly, or lifted from existing lattices element-wise or
point-wise.

>>>>

In this section, we simply let @R[T] = T@ (i.e., @NoRep@) to instantiate the
unstaged lattice type class. An simple instance of lattice we used is the power
set abstract domain (shown in Figure \ref{fig:powerset}), where the ordering is subset relation, the join is union
operation, and the meet is intersect operation.
Other lattices used in the remainder of the paper, such as products and maps, can be
implemented similarly, or lifted from existing lattices element-wise or
point-wise.

### 2019-03-29 paper/70_eval.tex
section 4
evaluation

\section{Evaluation} \label{evaluation}

To evaluate whether the performance can be improved through staging, we use the
abstract interpreter demonstrated in the paper to analyze some Scheme programs.
We first implement a desugaring transformation for a large subset of Scheme that
transforms into the small language we used in the paper. We use the same generic
interface with a 0-CFA-like abstraction, and the unstage abstract interpreter
forms the baseline.

All of our evaluation benchmarks were performed on an Ubuntu 16.04 LTS (Linux
kernel 4.4.0) machine with 4 Intel Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of
RAM. Although the machine has 96 cores and 192 threads in total, the abstract
interpreters only use one thread to run all the benchmark programs.

Our evaluations show that the staged abstract interpreter performs well in some
cases that are considered to be the worst scenarios for traditional $k$-CFA. The
staged version outperforms the unstaged abstract interpreter by a significant
margin (up to ~230\% performance gain) on these difficult ones.

\subsection{Benchmarks}

The benchmark programs we used are collected from previous papers
\cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
DBLP:journals/corr/abs-1102-3676} and exisiting artifacts
\footnote{https://github.com/ilyasergey/reachability}.

@@

\subsection{Performance}

Considering that our abstract interpreters are implemented in Scala which will be 
affected by the JIT warm up times, we run all experiments 10 times and report
the statistical median values of the running times.

\todo{remove fib; report LOC; mean}

\begin{figure}[h]
\begin{tabular}{@{}llll@{}}
\toprule
    Program            & unstaged   & staged     & $\Delta$ of median \\ \midrule
  % fib                & 1.020      & 2.046      & -50.14\%          \\
    rsa                & 35.928     & 145.306    & -75.27\%          \\
    kcfa3              & 2.934      & 5.671      & -48.27\%          \\
    church             & 884.800    & 545.736    & +62.13\%          \\
    fermat             & 15.673     & 68.790     & -77.22\%          \\
    kcfa-worst-16      & 320.689    & 301.564    & +6.34\%           \\
    kcfa-worst-32      & 3,685.615  & 1,854.461  & +98.74\%          \\
    kcfa-worst-64      & 49,422.666 & 14,849.725 & +232.82\%         \\
    \bottomrule
\end{tabular}
\caption{Evaluation result.} \label{evaluation_result}
\end{figure}

Figure~\ref{evaluation_result} shows the benchmark evaluation results. The
"unstaged" and "staged" columns show the median time to finish the analysis (in
milliseconds). The "$\Delta$ of median" column shows the performance change from
the unstaged abstract interpreter to the staged version.

As we can see from the table, the staged version significantly outperforms the
unstaged especially for the difficult ones. However, for some benchmarks, we
observe degraded performance. We conjecture this happens due to the fact that
the large size of the generated code is unoptimized -- as what we test here is a
simple implementation almost identical to what we described in the previous
sections, we believe the performance still can be improved if more engineering
effort is taken, such as optimizing code generation and avoiding code
duplication.

>>>>

\section{Empirical Evaluation} \label{evaluation}

To evaluate the performance can be improved by our staging abstract interpreters
approach, we scale our toy interpreter to a large subset of the Scheme language
and evaluate the analyzing time against with the unstaged versions. The result
shows an average \todo{X}\% running time improvement, as a strong evidence of
our \textit{abstraction without regret} approach to construct abstract
interpreters.

We implement the abstract interpreters for control-flow analysis, i.e., a
value-flow analysis focus on lambda terms in functional languages. A frontend
desugars Scheme to a small core language that make the analyzer easy to
implement. Macros and @call/cc@ is not handled. As described in Section ???, we
implement a monovariant 0-CFA-like analysis, a context-sensitive $1$-CFA-like
analysis, both their versions with and without store-widening.

\paragraph{Setting}

We use Scala 2.12.4 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04LTS
(kernel 4.4.0). All of our evaluation were performed on a machine with 4 Intel
Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96
cores and 192 threads in total, the abstract interpreters only use one thread.
To minimize the effect of warming-up from the JVM JIT compiler, all the
experiments are executed for 20 times and we report the statistical median value
of the running times (in seconds).

\paragraph{Benchmarks} \todo{Rewrite}
The benchmark programs we used are collected from previous papers
\cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
DBLP:journals/corr/abs-1102-3676} and exisiting artifacts
\footnote{https://github.com/ilyasergey/reachability}.

@@

\paragraph{Result}

\begin{figure}[h]
\tiny
\begin{tabular}{@{}lllll@{}}
\toprule
    Program          &\#AST & unstaged   & staged     & $\Delta$ \\ \midrule
  % fib              &      & 1.020      & 2.046      & -50.14\%          \\
    @rsa@            &      &            &            & \%          \\
    @kcfa3@          &      &            &            & \%          \\
    @church@         &      &            &            & \%          \\
    @fermat@         &      &            &            & \%          \\
    @kcfa-worst-16@  &  198 & 0.352      & 0.028      & +1157.14\%        \\
    @kcfa-worst-32@  &  390 & 4.234      & 0.289      & +1365.05\%          \\
    @kcfa-worst-64@  &  774 & 61.475     & 4.425      & +1289.27\%         \\
    \bottomrule
\end{tabular}
\caption{Evaluation result for 0CFA without store-widening.} \label{evaluation_result}
\end{figure}

Figure~\ref{evaluation_result} shows the evaluation results. The
\textit{unstaged} and \textit{staged} columns show the median time to finish the
analysis (in seconds). The "$\Delta$" column shows the median runtime change
from the unstaged to the staged version. \todo{0cfa, 1cfa, store-widen}

\iffalse
As we can see from the table, the staged version significantly outperforms the
unstaged especially for the difficult ones. However, for some benchmarks, we
observe degraded performance. We conjecture this happens due to the fact that
the large size of the generated code is unoptimized -- as what we test here is a
simple implementation almost identical to what we described in the previous
sections, we believe the performance still can be improved if more engineering
effort is taken, such as optimizing code generation and avoiding code
duplication.
\fi
